### 2019-01-27 writing/ehp_tables.tex

\begin{table}%[tbhp] 
\centering 
\caption{Exposure metrics considered to assess county-level exposure to 
tropical cyclones}
\begin{tabular}{p{1.5cm}p{2.5cm}p{10cm}} 
\toprule
Metric & Available & Criteria for exposure \\ \midrule 
Distance & 1988--2015 & County population mean center within 100 \si{\kilo\metre} of storm track \\ 
Rain & 1988--2011 & County cumulative rainfall of 75 millimeters or more over the period from two days before to one day after the storm's closest approach and county population mean center within 500 \si{\kilo\metre} of the storm track \\ 
Wind & 1988--2015 & Modeled maximum sustained wind speed at the county's population mean center 15 meters per second or higher during the storm\\ 
Flood & 1996--2015 & Flood event listed in the \ac{NOAA} Storm Events database for the county with a start date within two days of the storm's closest approach and county population mean center within 500 \si{\kilo\metre} of the storm track \\
Tornado & 1996--2015 & Tornado event listed in the \ac{NOAA} Storm Events database for the county with a start date within two days of the storm's closest approach and county population mean center within 500 \si{\kilo\metre} of the storm track\\
\bottomrule 
\end{tabular} 
\label{tab:exposuremetrics} 
\end{table}

\clearpage

\input{../tables/stormseverity.tex}