### 2019-01-28 writing/anderson_ehp_supp.tex

\documentclass[11pt, titlepage, twoside]{article}

\input{ehp_preamble}

\begin{document}

\title{Supplemental Material for ``Assessing United States county-level
exposure for research on tropical cyclones and human health"}

\author[1,*]{G. Brooke Anderson} 
\author[1,2]{Joshua Ferreri}
\author[3]{Mohammad Al-Hamdan} 
\author[3]{William Crosson} 
\author[4]{Andrea Schumacher} 
\author[5]{Seth Guikema} 
\author[6]{Steven Quiring} 
\author[7]{Dirk Eddelbuettel} 
\author[1]{Meilin Yan} 
\author[8]{Roger D. Peng}

\affil[1]{Department of Environmental \& Radiological Health Sciences, Colorado 
  State University, Fort Collins, CO, 80523} 
\affil[2]{University of Colorado Denver School of Medicine, Aurora, CO, 80045} 
\affil[3]{Universities Space Research Association, NASA Marshall Space Flight 
  Center, Huntsville, AL, 35805}
\affil[4]{Cooperative Institute for Research in the Atmosphere, Colorado State
  University, Fort Collins, CO, 80523} 
\affil[5]{Department of Industrial and Operations Engineering, University of 
  Michigan, Ann Arbor, MI, 48109}
\affil[6]{Department of Geography, Ohio State University, Columbus, OH, 43210}
\affil[7]{Debian and R Projects; Department of Statistics, University of
  Illinois at Urbana-Champaign, Champaign, IL, 61820} 
\affil[8]{Department of Biostatistics, Johns Hopkins Bloomberg School of Public 
  Health, Baltimore, MD, 21205}
\affil[*]{Corresponding author: Brooke Anderson, brooke.anderson@colostate.edu}

\date{}

\maketitle

\beginsupplement

\begin{figure}[tbhp!]
\centering
\includegraphics[width=\linewidth]{previous_exposure_metrics}
\caption{Metrics used in a sample of previous studies to assign exposure to
tropical cyclones in \ac{US}-based tropical cyclone exposure or impact studies.
Citations for each study are shown on the left and the broad topic of the study
on the left. Presence of a bar for a study-metric combination indicates that
the given metric was used in the study in assessing exposure to tropical
cyclones for that study, either alone or in combination with other metrics for
which that study has bars. The color transparency of each bar indicates the
number of different thresholds considered for defining storm exposure in the
study under that hazard (e.g., less transparent bars indicate that a study
considered multiple thresholds of the metric in defining tropical cyclone
exposure as a method of sensitivity analysis). Studies for which there are two
or more bars (e.g., bars for both distance and wind) indicate studies that
either used separate exposure assessments, one for each metric, (e.g., as a
sensitivity analysis) or used a single exposure index that incorporated
multiple metrics in its definition. ``Damage" often indicate use of Federal
Emergency Management Agency (FEMA) reports, although occasionally used damage
reports from other sources, including newspaper reports. Examples of ``Other"
include evacuation orders, excessive school closures, length of time a storm
was near a location, and detailed, holistic analyses of a specific storm or
storms. References for each study are included in the bibliography of this
Supplemental Material.}
\label{fig:previousmetrics}
\end{figure}

\nocite{logan2015, gares2014, esnard2011, zandbergen2009, horney2016,
grabich2016, antipova2015post, zahran2010, currie2013, grabich2015,
grabich2016hurricane, amstadter2010, acierno2006, le2013, lieberman2017,
mckinney2011, dosa2012evacuate, swerdel2014, czajkowski2011, mongin2017,
kessler2007hurricane, caillouet2008increase, rosenheim2018disaster,
domino2003disasters, kim2016, kinney2008, kim2015, tansel2010, escobedo2009,
fuller2014, baggerly2008impact, strobl2011economic, mukherjee2017}

\newpage

\begin{figure}[tbhp!]
\centering
\includegraphics[width=0.7\linewidth]{hurrtracks}
\caption{Tracks of all tropical cyclones included in the study. The study
included all tracked storms in the revised Atlantic hurricane database
\ac{HURDAT2}~\parencite{landsea2013} that came within 250~\si{\kilo\metre} of at least on
\ac{US} county, 1988\,--\,2015. Thicker lines show the tracks of storms whose
names were retired, indicating that the storm was particularly severe or had
notable impacts~\parencite{retirednames}.}
\label{fig:hurrtracks}
\end{figure}

\clearpage

\begin{figure}[tbhp!]
\centering
\includegraphics[width=\linewidth]{othertopstorms}
\caption{Counties exposed, based on different exposure metrics, to the storms
with largest extent (as measured by the number of counties exposed based on any
metric) from each of the clusters shown in the Jaccard heatmap in the main text
(Figure 2\todo[inline]{Check fig ref}).}
\label{fig:othertopstorms}
\end{figure}

\clearpage

\begin{figure}[tbhp!]
\centering
\includegraphics[width=\linewidth]{extentdisagreement}
\caption{Examples of storms for which exposure assessments differed
substantially between rain-based and wind-based metrics. Counties shown in
orange were classified as exposed to a given storm by a given metric; those in
yellow were classified as not exposed. The storm's track is shown in red. For
both Frances (2004) and Isidore (2002) (top panel), substantially more counties
were assessed as exposed when measuring exposure based on rain rather than
wind. Conversely, for both Hugo (1989) and Isabel (2003) (bottom panel),
substantially more counties were assessed as exposed based on wind compared to
rain.}
\label{fig:extentdisagreement}
\end{figure}

\clearpage

\begin{figure}[tbhp!]
\centering
\includegraphics[width=\linewidth]{averageexposureonlysupp}
\caption{Average number of storm exposure per decade in \ac{US} counties for
each exposure metric, based on years of for which data for all five exposure
metrics are available (1996\,--\,2011). The criteria behind each of the five
metrics is given in Table 1 of the main text.} \label{fig:averageexposuresupp}
\end{figure}

\clearpage

\begin{figure}[tbhp!]
\centering
\includegraphics[width=\linewidth]{windexamples}
\caption{Wind classifications for the two storms, Hurricanes Sandy (2012) and
Ike (2008) for which exposure classifications from modeled wind
data~\parencite{stormwindmodel} and \ac{HURDAT2} wind
radii~\parencite{demuth2006improvement} showed the highest disagreement for storms
between 2004 and 2015 (Figure~\ref{fig:windcomparison}). The top two maps show
modeled (left) and \ac{HURDAT2} (right) windspeed categorizations for Hurricane
Ike (2008), while the bottom two maps show the same windspeed characterizations
for Hurricane Sandy (2012).  In both cases, the storms were exceptionally large
systems for which high winds persisted well inland from landfall and for which,
based on the \ac{HURDAT2} data, a large number of counties were exposed to
winds of $\ge$34 knots (Figure~\ref{fig:windcomparison}\todo[inline]{Fix fig ref}).}
\label{fig:windexamples}
\end{figure}

\begin{figure}[tbhp!]
\centering
\includegraphics[width=0.9\linewidth]{rainexamples}
\caption{Examples of storms where the rain exposure classification method fails
to capture full storm-related rainfall because of the distance constraint used
in that assessment metric. The red line on each map shows the track of the
storm, based on the revised Atlantic hurricane database
\ac{HURDAT2}~\parencite{landsea2013}. The color of each county in the map gives the
cumulative precipitation (in \si{\milli\metre}) for that county during the window
from two days before to one day after the storm's closest approach to the
county. The blue outline identifies the collection of counties that were
classified as ``exposed" based on the criteria that the storm passed within
500~\si{\kilo\metre} of the county's population mean center and the cumulative
rainfall was 75 \si{\milli\metre} or more (Table 1 of main text). The map on the left
shows rain-based exposure for Hurricane Ike. For this exceptionally large
storm, there are some counties in northern Texas that received high rainfall
from the storm but are too far from the storm center's track to be classified
as exposed by this metric. The map on the right shows rain-based exposure for
Tropical Storm Lee. For this storm, rain from the storm affected counties in
the Mid-Atlantic and New England as the remnants of the storm moved over those
regions, but storm tracking stopped while the storm was still over the
southeastern part of the \ac{US}, and so these counties are not classified as
exposed to the storm based on this metric.}
\label{fig:rainexamples}
\end{figure}

\clearpage

\printbibliography

\cleardoublepage

\listoftodos

\end{document}