### 2019-07-27 writing/ehp_figures.tex

\section*{Figures}

\listoffigures

\clearpage

\begin{figure}[tbhp!] \centering
\includegraphics[width=0.7\linewidth]{hurrtracks}
\caption{States and storms considered in this study. All counties in the states
        shown in this map were investigated. The lines show the paths
	of the study storms, which included all tracked storms in 1988\,--\,2015
	\ac{HURDAT2}~\parencite{landsea2013} that came within 250~\si{\kilo\metre} 
	of at least one \ac{US} county. Thicker lines show the tracks 
	of storms whose names have been retired, indicating that the storm was particularly 
	severe or had notable impacts~\parencite{retirednames}.
	}
\label{fig:hurrtracks}
\end{figure}

\clearpage

\begin{figure}[tbhp!] \centering
\includegraphics[width=0.9\linewidth]{raincomparison} 
	\caption{Comparison of storm-associated rainfall estimates between 
	a re-analysis dataset (included as the rainfall exposure metric
	in the open-source data and included in further analysis) and ground-based
	observations in nine sample counties. The rainfall estimates  
	include rainfall from two days before to one day after the storm's closest
	approach to the county. Each small graph shows data for one sample county, 
	and each point shows one tropical storm. The number of storms 
	within each county and the average number of stations reporting rainfall during 
	the county's storms are given above each plot. Horizontal and vertical lines 
	in each plot show the threshold of 75~\si{\milli\metre} used to classify a storm 
	as ``exposed'' in further analysis (Table 1). Note that 
	x- and y-axis ranges differ by county.
	} 
\label{fig:raincomparison}
\end{figure}

\begin{figure}[tbhp!]
\centering
\includegraphics[width=0.6\linewidth]{windcomparison}
\caption{Comparison of modeled maximum sustained surface wind speeds 
	(included as the primary wind exposure metric in the open-source
	data and used in further analysis) with estimates from the \ac{HURDAT2}
	wind speed radii. Each point represents a storm, with the x-axis
	giving the percent of counties classified in the same category
	of maximum sustained surface wind speeds ($<$34~\si{\knot}; 
	34\,--\,49.9~\si{\knot}; 50\,--\,63.9~\si{\knot}; 
	$\ge$64~\si{\knot}) by both sets of wind data for that storm. The 
	color of each point indicates the extent of the storm, 
	giving the number of study counties that were exposed to 
	maximum sustained surface wind speeds of at least 34~\si{\knot}
	(based on modeled wind speeds). Estimates are shown for [all?]
	study storms since 2004, the earliest year for which 
	wind speed radii estimates are available in \ac{HURDAT2} [?].
	}
\label{fig:windcomparison}
\end{figure}

\begin{figure}[tbhp!]
\centering
\includegraphics[width=0.8\linewidth]{floodcomparison}
\caption{Comparison for a sample of study counties of NOAA Storm Events
listings versus five-day total streamflows at county streamflow gages during
tropical storms. The number of streamflow gages used in analysis are given in
parentheses beside the county's name in the panel title. Each point represents
a single tropical storm. Note that the x-axis scales differ by county,
depending on the number of streamflow gages and typical flow rates for each
gage, and are on a log-10 scale.}
\label{fig:floodcomparison}
\end{figure}

\clearpage

\begin{figure}%[tbhp] 
\centering
\includegraphics[width=15cm]{averageexposureonly.pdf} 
\caption{Average number of storm exposures per decade in U.S. counties for each
exposure metric. The criteria behind each of the five metrics is given in Table
\ref{tab:exposuremetrics}. The years used to estimate these averages are based
on years of available exposure data (distance and wind:~1988\,--\,2015; rain:
1988\,--\,2011; flood and tornado:~1996\,--\,2015). } 
\label{fig:averageexposure} 
\end{figure}

\clearpage

\begin{figure}%[tbhp]
\centering
\includegraphics[width=16cm]{ivanonly.pdf}
\caption{Counties classified as exposed to Hurricane Ivan in 2004 under each
exposure metric considered (Table~\ref{tab:exposuremetrics}). The red line
shows the track of Hurricane Ivan based on the revised Atlantic hurricane
database \ac{HURDAT2}~\parencite{landsea2013}.  Similar maps for other
large-extent storms are given in Figure~S3.}
\label{fig:ivanexposure} 
\end{figure}

\clearpage

\begin{figure}%[tbhp] 
\centering 
\includegraphics[width = 0.6\linewidth]{jaccard_heatmap.pdf} 
\caption{Agreement between exposure classifications based on different
         exposure metrics for all storms between~1996 and~2011 for which 
	 at least~250 counties were exposed based on at least one metric.
	 Each row shows one storm, and the color of each cell shows the 
	 measured Jaccard index for each pair of exposure metrics 
	 (proportion of counties classified as exposed by both metrics out 
	 of storms classified as exposed by either metric). The colors to the 
	 right of the main heatmap for each storm indicate the total number of 
	 counties classified as exposed to the storm by any of the five metrics, 
	 providing an indication of each storm's extent. Storms are displayed 
	 within clusters that have similar patterns in county-level exposure 
	 agreement for metric pairs, based on hierarchical clustering using the 
	 complete link method~\parencite{murtagh2012algorithms}; columns are also 
	 ordered based on hierarchical clustering. Maps are available showing the 
	 counties identified as exposed under each of five metrics for the widest-extent 
	 storm in each cluster: Hurricane Ivan in~2004 (Figure~\ref{fig:ivanexposure}) 
	 and Hurricanes Floyd in~1999, Lee in~2011, Cindy in~2005, and Katrina 
	 in~2005 (Figure~S3).
} 
\label{fig:jaccard}
\end{figure}

\clearpage

\begin{figure*}%[tbhp]
\centering
\includegraphics[width=15.5cm]{topelecdependexposure}
\caption{Counties with the highest estimated physical exposure among
electricity-dependent Medicare beneficiaries exposed to storms per year in U.S.
counties for each exposure metric. The color of each bar indicates the number
of Medicare beneficiaries in the county reliant on electricity-dependent
medical and assistive equipment as of July~2017. The length
of each bar shows the average expected number of these electricity-dependent
Medicare beneficiaries exposed to tropical storms per year based on a given
exposure metric.}
\label{fig:topelecdependexposure}
\end{figure*}