### 2019-07-28 tables/stormseverity.tex
Update storm severity table

% latex table generated in R 3.4.3 by xtable 1.8-2 package
% Wed Oct 10 12:11:35 2018
\begin{table}[ht]
\centering
\caption{Summary statistics for the number of county tropical cyclone exposures under each metric.     
	 The median and interquartile range of number of county exposures per tropical cyclone are based on
         the tropical cyclones for which at least one US county was exposed. The years for which
         data are available for each metric are given in Table 1.} 
\label{tab:exposuresummaries}
\begin{tabular}{p{1.75cm}p{3cm}p{4cm}p{4cm}}
  \toprule
Metric & Mean (interquartile range) of county exposures per year & Median (interquartile range) of county exposures per tropical cyclone & Tropical cyclone with most counties exposures (\# exposed counties) \\ 
  \midrule
Distance & 433 (224, 503) & 84 (23, 173) & Beryl, 1994 (330) \\ 
  Rain & 401 (152, 553) & 68 (14, 146) & Frances, 2004 (464) \\ 
  Wind & 224 (83, 353) & 34 (8, 80) & Ike, 2008 (355) \\ 
  Flood & 213 (84, 249) & 26 (6, 57) & Ivan, 2004 (317) \\ 
  Tornado & 53 (16, 43) & 7 (2, 16) & Ivan, 2004 (91) \\ 
   \bottomrule
\end{tabular}
\end{table}