### 2019-10-09 writing/ehp_tables.tex

\begin{table}%[tbhp] 
\centering 
\caption{Exposure metrics considered to assess county-level exposure to 
tropical cyclones}
\begin{tabular}{p{1.5cm}p{2.5cm}p{10cm}} 
\toprule
Metric & Available & Criteria for exposure \\ \midrule 
Distance & 1988\,--\,2015 & Storm track came within 100~\si{\kilo\metre} of county population mean center\\ 
Rain & 1988\,--\,2011 & County cumulative rainfall of $\ge$75~\si{\milli\metre} over the period from two days before to 
one day after the storm's closest approach and the storm track came within 500~\si{\kilo\metre} of the county population mean center\\ 
Wind & 1988\,--\,2015 & Modeled storm-associated maximum sustained winds of $\ge$34~kts 
at the county's population mean center\\ 
Flood & 1996\,--\,2015 & Flood event listed in the \ac{NOAA} Storm Events database for the county with a start date within two days of the storm's closest approach and the storm track came within 500~\si{\kilo\metre} of the county population mean center \\
Tornado & 1996\,--\,2015 & Tornado event listed in the \ac{NOAA} Storm Events database for the county with a start date within two days of the storm's closest approach and the storm track came within 500~\si{\kilo\metre} of the county population mean center\\
\bottomrule 
\end{tabular} 
\label{tab:exposuremetrics} 
\end{table}

\clearpage

\input{../tables/stormseverity.tex}