### 2019-11-07 writing/ehp_introduction.tex

\section*{Introduction}

\acresetall

Recently, several major tropical cyclones have hit \ac{US} cities, including
Hurricanes Harvey, Irma, and Maria in 2017~\parencite{blake20182017} and
Hurricanes Florence and Michael in 2018 \parencite{avila20192018}. Researchers
are exploring how these storms impacted human health
(e.g.,~\cite{santos2018use, rivera2018estimating, santos2018differential,
grineski2019impact, issa2018deaths, tanz2019notes, paul2019brief}), including
through projects funded by \ac{NIH} Rapid Response
grants~\parencite{nihreporter}. These projects will add to evidence
characterizing how hurricanes impact human health, including evidence from
multi-year studies based on fatality data from post-disaster
surveillance~\parencite{rappaport2000, rappaport2014fatalities,
rappaport2016fatalities, czajkowski2010fatal, czajkowski2011, moore2012}.

For several types of climate-related disasters---including heat waves, floods,
and wildfires---research based on post-disaster surveillance data has been
richly supplemented by studies analyzing trends in community-wide
administrative or vital data, including time series or case-crossover studies
(e.g.,~\cite{anderson2010heat, son2012impact, haikerwal2015impact,
liu2017wildfire, milojevic2017mental}).  These studies typically compare the
rate of a health outcome in a community during a disaster or disasters to rates
during comparable non-disaster periods.  While these studies cannot attribute
specific cases (e.g., a specific person's death) to a disaster, they can help
characterize the community-wide change in health risk during disasters, and can
help capture impacts that might be missed or underestimated in post-disaster
surveillance.  In many cases, these studies analyze multi-year, multi-community
data, allowing them to estimate average associations over many disasters and to
explore how a disaster's characteristics, or the characteristics of the
community it hits, modify associated health risk (e.g.,~\cite{anderson2010heat,
son2012impact, liu2017wildfire}).  Some studies have begun to use this approach
to study the health impacts of tropical cyclones, including several studies of
Hurricane Maria (e.g.,~\cite{santos2018use, santos2018differential}), Hurricane
Sandy (e.g.,~\cite{kim2016, mongin2017, swerdel2014}), and the 2004 hurricane
season in Florida~\parencite{mckinney2011}.  However, to expand this approach
to study longer time periods for large sets of communities, researchers must be
able to assess exposure to tropical cyclones consistently and comparably across
storms, years, and communities.

The \ac{NHC} publishes a ``Best Tracks'' dataset that is considered the gold
standard for Atlantic-basin tropical cyclone tracks.  It records the central
position of a tropical cyclone every six hours, as well as the storm's minimum
central pressure and maximum wind speed.  This data is openly available through
the \ac{HURDAT2}, a post-storm assessment that incorporates data from several
sources, including satellite data and, when available, aircraft reconnaissance
data~\parencite{landsea2013, jarvinen1988}.  With this data, it is
straightforward to measure whether a community was in the direct path, or
within a certain distance of the path, of a storm. Some studies have indeed
used this data within epidemiological studies to assess tropical cyclone
exposure based on whether a storm's track passed through a community or how
near the storm's central track came to the community (e.g.,~\cite{currie2013,
kinney2008, caillouet2008increase}).

When epidemiological studies assess exposure to a tropical cyclone using, as a
proxy measurement, how close the storm came to a community, they may
misclassify exposure~\parencite{grabich2015measuring}. While a number of
tropical cyclone hazards are strongly associated with distance from the
tropical cyclone's center (e.g., wind and, at the coast, storm surge and
waves~\parencite{rappaport2000, kruk2010}), other hazards like heavy rainfall,
floods, and tornadoes can occur well away from the tropical cyclone's central
track~\parencite{rappaport2000, atallah2007, moore2012}.  For example, fatal
tropical cyclone tornadoes, which were linked to over~300 deaths in the \ac{US}
between~1995 and~2009, most often occur~200\,--\,500~\si{\kilo\metre} from the
tropical cyclone's center~\parencite{moore2012}.  Further, when studies use
distance from the storm's track to assess exposure, they often use the same
buffer constraints on each side of the tropical cyclone track
(e.g.,~\cite{czajkowski2011, kinney2008, currie2013}).  However, the forces of
a tropical cyclone tend to be distributed around its center non-symmetrically.
Extreme winds are more common to the track's right, where counter-clockwise
cyclonic winds move in concert with the tropical cyclone's forward
motion~\parencite{halverson2015}, and almost all of the fatal tornadoes
associated with \ac{US} tropical cyclones between~1995 and~2009 occurred to the
right of the tropical cyclone's track, mostly in the storm's right front
quadrant~\parencite{moore2012}. Rain, conversely, is often heaviest to the left
of the track, especially when the tropical cyclone interacts with other weather
systems~\parencite{atallah2003, atallah2007, zhu2013variations} or undergoes an
extratropical transition~\parencite{elsberry2002}.

The multi-hazard nature of tropical cyclones therefore makes it hard to assess
exposure based on how close the storm's central
track came to the community.  While other approaches have been developed to
incorporate storm hazards, particulary wind, into exposure assessment (e.g.,
\cite{grabich2015measuring, zandbergen2009, czajkowski2011}), there is not yet
a standard approach, and when different studies use different datasets or storm
hazards when assessing storm exposure, it becomes difficult to compare and
aggregate findings from different studies.

To help epidemiological researchers assess exposure to tropical cyclones, we
developed an open-source dataset~\parencite{hurricaneexposure}.  These data
cover all counties in states in the eastern half of the \ac{US} over multiple
years and include five metrics characterizing exposure to tropical cyclones
(Table~\ref{tab:exposuremetrics}): (1)~closest distance the storm's central
track came to the county's center (a proxy measurement of storm exposure used
in some previous studies); (2)~maximum sustained wind speed at the county's
center over the course of the storm; (3)~cumulative rainfall in the county over
the course of the storm; (4)~flood events in the county associated with the
storm; and (5)~tornado events in the county associated with the storm.  We
aggregated these data at the county level since data on many potential impacts
are available at county-level aggregations (e.g., direct hurricane-related
deaths~\parencite{czajkowski2011}; birth outcomes~\parencite{grabich2015,
grabich2015measuring}; autism prevalence~\parencite{kinney2008}) and since
decisions and policies to prepare for, and respond to, tropical cyclones are
often undertaken at the county level~\parencite{zandbergen2009, rappaport2000}.
We also wrote functions to explore and map this data and to link it
with county-level health data~\parencite{hurricaneexposuredata}

Here, we describe how we developed this dataset and explore how its
measurements compare to other data that could be used to characterize tropical
cyclone hazards in the context of epidemiological research. \textit{Further, we expand
on previous research that found large differences in which communities are
assessed as exposed to a tropical cyclone depending on whether exposure
assessment is based on distance to the storm's track or if it uses a metric
incorporating storm-related winds within the county during the storm,
concluding that a study may be prone to bias from exposure misclassification if
distance to the storm track is used as a proxy measurement for exposure
assessment~\parencite{grabich2015measuring}.}  The previous study investigated
this question for one hurricane season in Florida. Here we expand to
investigate this question across all counties in states in the eastern half of
the \ac{US} over multiple storms and years.  Further, we investigate patterns
in exposure classification based on other metrics of storm-related
hazards---rainfall, flooding, and tornadoes---that are important to characterize to
improve understanding of the inland health impacts of tropical
cyclones~\parencite{czajkowski2011, moore2012}. These results can help
epidemiologists design studies and plan statistical analysis for
multi-year, multi-community studies that investigate changes in community-wide
rates of health outcomes during exposure to the hazards brought by tropical
cyclones, including counties inland from a storm's landfall.