### 2019-11-12 writing/ehp_introduction.tex

\section*{Introduction}

\acresetall

Hurricanes and other tropical cyclones can severely impact human health in
\ac{US} communities, and some of these impacts have been characterized through
multi-year studies based on fatality data from post-disaster
surveillance~\parencite{rappaport2000, rappaport2014fatalities,
rappaport2016fatalities, czajkowski2010fatal, czajkowski2011, moore2012}.  For
several types of climate-related disasters---including heat waves, floods, and
wildfires---research based on post-disaster surveillance data has been richly
supplemented by studies analyzing trends in community-wide administrative or
vital data, including time series or case-crossover studies
(e.g.,~\cite{anderson2010heat, son2012impact, haikerwal2015impact,
liu2017wildfire, milojevic2017mental}).  These studies typically compare the
rate of a health outcome in a community during a disaster to rates during
comparable non-disaster periods.  While these studies cannot attribute specific
cases (e.g., a specific person's death) to a disaster, they can help
characterize the community-wide change in health risk during disasters and
capture impacts that might be missed or underestimated in post-disaster
surveillance.  In many cases, these studies analyze multi-year, multi-community
data, allowing them to estimate average associations over many disasters and to
explore how a disaster's characteristics, or the characteristics of the
community it hits, modify associated health risks (e.g.,~\cite{anderson2010heat,
son2012impact, liu2017wildfire}).  Some studies have begun to use this approach
to study the health impacts of tropical cyclones, including several studies of
Hurricane Maria (e.g.,~\cite{santos2018use, santos2018differential}), Hurricane
Sandy (e.g.,~\cite{kim2016, mongin2017, swerdel2014}), and the 2004 hurricane
season in Florida~\parencite{mckinney2011}.  However, to expand this approach
to study longer time periods for large sets of communities, researchers must be
able to assess exposure to tropical cyclones consistently and comparably across
storms, years, and communities.

The \ac{NHC} publishes a ``Best Tracks'' dataset that is considered the gold
standard for Atlantic-basin tropical cyclone tracks.  It records the central
position of a tropical cyclone every six hours, as well as the storm's minimum
central pressure and maximum wind speed.  These data are openly available
through the \ac{HURDAT2}, a post-storm assessment that incorporates data from
several sources, including satellite data and, when available, aircraft
reconnaissance data~\parencite{landsea2013, jarvinen1988}.  With these data, it
is straightforward to measure whether a community was in the direct path, or
within a certain distance of the central track, of a storm, as some previous
epidemiological studies have done to assess exposure to a tropical storm
(e.g.,~\cite{currie2013, kinney2008, caillouet2008increase}).

However, when epidemiological studies assess exposure to a tropical cyclone
using, as a proxy measurement, how close the storm came to a community, they
may misclassify exposure~\parencite{grabich2015measuring}. While a number of
tropical cyclone hazards are strongly associated with distance from the
tropical cyclone's center (e.g., wind and, at the coast, storm surge and
waves~\parencite{rappaport2000, kruk2010}), other hazards like heavy rainfall,
floods, and tornadoes can occur well away from the tropical cyclone's central
track~\parencite{rappaport2000, atallah2007, moore2012}.  For example, fatal
tropical cyclone tornadoes, which were linked to over~300 deaths in the \ac{US}
between~1995 and~2009, most often occur~200\,--\,500~\si{\kilo\metre} from the
tropical cyclone's center~\parencite{moore2012}.  Further, when studies use
distance from the storm's track to assess exposure, they often use the same
buffer constraints on each side of the storm track
(e.g.,~\cite{kinney2008, currie2013}).  However, the forces of a tropical
cyclone tend to be distributed around its center non-symmetrically.  Extreme
winds are more common to the track's right, where counter-clockwise cyclonic
winds move in concert with the tropical cyclone's forward
motion~\parencite{halverson2015}, and almost all of the fatal tornadoes
associated with \ac{US} tropical cyclones between~1995 and~2009 occurred to the
right of the tropical cyclone's track~\parencite{moore2012}. Rain, conversely,
is often heaviest to the left of the track, especially when the tropical
cyclone interacts with other weather systems~\parencite{atallah2003,
atallah2007, zhu2013variations} or undergoes an extratropical
transition~\parencite{elsberry2002}.  The multi-hazard nature of tropical
cyclones therefore makes it hard to assess exposure based on how close the
storm's central track came to the community.  While other approaches have been
developed to incorporate storm hazards, particulary wind, into exposure
assessment (e.g., \cite{grabich2015measuring, zandbergen2009, czajkowski2011}),
there is not yet a standard approach, and when different studies use different
datasets or storm hazards when assessing storm exposure, it becomes difficult
to compare and aggregate findings from different studies.

To help epidemiological researchers assess exposure to tropical cyclones, we
developed an open-source dataset~\parencite{hurricaneexposure}.  These data
cover all counties in states in the eastern half of the \ac{US} over multiple
years and include five metrics characterizing exposure to tropical cyclones
(Table~\ref{tab:exposuremetrics}): (1)~closest distance the storm's central
track came to the county's center (a proxy measurement of storm exposure used
in some previous studies); (2)~maximum sustained wind speed at the county's
center over the course of the storm; (3)~cumulative rainfall in the county over
the course of the storm; (4)~flooding in the county
concurrent with the storm; and (5)~tornadoes in the county
concurrent with the storm.  We aggregated these data at the county level since
data for epidemiological studies are often available at county-level
aggregations (e.g., direct hurricane-related deaths~\parencite{czajkowski2011};
birth outcomes~\parencite{grabich2015, grabich2015measuring}; autism
prevalence~\parencite{kinney2008}) and since decisions and policies to prepare
for, and respond to, tropical cyclones are often undertaken at the county
level~\parencite{zandbergen2009, rappaport2000}.  We also wrote functions to
explore and map these data and to link it with county-level health
data~\parencite{hurricaneexposuredata}

Here, we describe how we developed this dataset and explore how its
measurements compare to other data that could be used to characterize tropical
cyclone hazards for epidemiological research. Further, we expand
on previous research that investigated methods to measure exposure to tropical
cyclones for epidemiological research~\parencite{grabich2015measuring}.  The
previous study investigated differences in which communities were assessed as
exposed to tropical cyclones during the 2004 hurricane season in Florida,
depending on whether exposure assessment was based on distance to the storm's
track or used a metric incorporating storm-related winds within the county
during the storm.  They found important differences across methods of assessing
storm exposure, concluding that a study may be prone to bias from exposure
misclassification if distance to the storm track is used as a proxy measurement
for exposure assessment~\parencite{grabich2015measuring}.  Here we expand to
investigate this question across a large set of counties and storm seasons.
Further, we investigate patterns in exposure classification based on other
metrics of storm-related hazards---rainfall, flooding, and tornadoes---that are
important to characterize to improve understanding of the inland health impacts
of tropical cyclones~\parencite{czajkowski2011, moore2012}. These results can
help epidemiologists design studies and plan statistical analysis for
multi-year, multi-community studies that investigate changes in community-wide
rates of health outcomes during exposure to the hazards brought by tropical
cyclones, including counties inland from a storm's landfall.