### 2019-11-12 writing/ehp_abstract.tex
Work some more on abstract and small wording

\begin{abstract}

\noindent \textbf{Background:} 
        Tropical cyclone epidemiology can be 
	advanced through county-level exposure assessment methods that are comprehensive and 
	consistent across space and time, allowing multi-year, multi-storm studies. 
	Further, an understanding of 
	patterns in and between hazard-specific exposure metrics can help in 
	designing tropical cyclone epidemiological studies.\\ 
\textbf{Objectives:} (1)~Provide an open-source dataset for tropical
	cyclone exposure assessment for epidemiological research and
	(2)~investigate patterns and agreement between county-level assessments of tropical cyclone
	exposure based on different storm hazards. \\ 
\textbf{Methods:} We created an open-source dataset with county-level data on
	exposure to four tropical cyclone hazards: maximum sustained winds,
	rainfall, flooding, and tornadoes. The data cover all eastern \ac{US} counties 
	for all land-falling or near-land
	Atlantic basin storms for~1996\,--\,2011 for all
	metrics and up to~1988\,--\,2015 for specific metrics. 	
	We validated 
	measurements against other data sources and investigated patterns
	and agreement among binary exposure classifications based on these
	metrics, as well as between distance from the storm's track as a proxy for
	exposure.\\ 
\textbf{Results:} Measurements in the dataset created here typically
        agreed with data from other sources, and we present and discuss areas of
	disagreement and other caveats for the data.
        Long-term geographic patterns in exposure classification agree with previous results from
	atmospheric science. County-level exposure classification for a storm differs substantially 
	depending on the hazard metric used. \\ 
\textbf{Discussion:} Our results provide a multi-hazard dataset that can be
	leveraged for epidemiologic research on tropical cyclones, as well as 
	insights that can inform the design of tropical cyclone epidemiological 
	studies.
\end{abstract}