### 2019-11-28 writing/ehp_results.tex

\section*{Results}

\subsection*{Exposure assessment data, data validation, and software}

We created and published this tropical cyclone exposure data as an R
package~\parencite{hurricaneexposuredata}.  The package size exceeded the
recommended maximum size for \ac{CRAN}, the standard repository for publishing
R packages. Therefore, we used the \textit{drat} framework to set up our own
repository to host the package~\parencite{anderson2017hosting}. These data
include continuous county-level measurements for the closest distance of each
storm, cumulative rainfall, and peak sustained surface wind. These continuous
metrics can be used for classifying counties as exposed or unexposed---using
thresholds selected by the user---or can be extracted directly as continuous
metrics. The dataset also include binary data on flood and tornado events
associated with each storm in each county. To accompany these data, we
published an additional R package with software tools to explore and map the
data and to integrate it with human health
datasets~\parencite{hurricaneexposure}.

We explored potential limitations in these data by comparing them with data
from other available sources.  For estimates of storm-associated rainfall, we
compared estimates included in the open-source data with ground-based
observations in nine sample counties (Figure~\ref{fig:raincomparison}). Within
these counties, storm-related rainfall measurements were well-correlated
between the two data sources, with rank correlations (bottom right of each
graph in Figure~\ref{fig:raincomparison}) between 0.87 and 0.98. There was some
evidence that the primary rainfall metric may tend to underestimate rainfall
totals in storms with extremely high rainfall, based on a few heavy-rainfall
storms in Harris County, TX, Mobile County, AL, Charleston County, SC, and Wake
County, NC (Figure~\ref{fig:raincomparison}). However, even in these cases, it
was rare for a storm to be classified differently (under the precipitation
threshold of 75~\si{\milli\metre} we used for binary exposure classification
for later analysis) based on the source of precipitation data---horizontal and
vertical lines in each small plot in Figure~\ref{fig:raincomparison} show the
threshold of 75~\si{\milli\metre}, so storms in the lower left and upper right
quadrants would be classified the same (``exposed'' or ``unexposed'')
regardless of the precipitation data source, while storms in the upper left and
lower right quadrants would be classified differently. Such cases were rare.

For peak sustained surface wind estimates, we found that the primary wind
exposure metric in the open-source data generally agreed well with the wind
radii reported in \ac{HURDAT2}. For most storms, $\ge$95\% of counties were
assigned the same category of wind speed ($<$34~kt; 34\,--\,49.9~kt;
50\,--\,63.9~kt; $\ge$64~kt) by both data sources
(Figure~\ref{fig:windcomparison}).  Disagreement was limited to a few storms
(e.g., Hurricanes Sandy in 2012 and Ike in 2008). For these two storms, the
modeled wind speed values used in the open-source data somewhat overestimated
the extent of the storm's most severe winds at landfall but then underestimated
the extent of 34\,--\,49.9 kt storm winds further inland (Figure~S2). For
epidemiology researchers who would like to conduct sensitivity analysis between
these two sets of wind data, we have included estimates from the \ac{HURDAT2}
wind radii as a secondary measure of county-level peak sustained wind in the
open-source dataset \parencite{hurricaneexposuredata}.

For flood data, we compared the flood status values included in the open-source
data to streamflow measurements at \ac{US} Geological Survey gages within nine
sample counties (Figure~\ref{fig:floodcomparison}). Across the sampled
counties, streamgage data generally indicated higher discharge during periods
identified as flood events based on the NOAA Storm Events database,
particularly in counties with more reporting streamgages (e.g., Harris County,
TX, with~10 reporting stream gages; Figure~\ref{fig:floodcomparison}). There
were some cases, however, where the two flooding data sources were somewhat
inconsistent.  For example, there were one or two tropical cyclones in several
of the counties (Mobile County, AL, Escambia County, FL, and Montgomery County,
PA) with associated flood event listings but for which the total discharge
across county streamflow gages was relatively low. For storms without a flood
listing for the county, in most cases the total streamflow discharge in the
county was relatively low, and, in all but one of these cases all streamgage
flows were below the flooding threshold. The exception was for Hurricane Ida in
2009 in Fulton County, GA.

\subsection*{Patterns in tropical cyclone exposures}

We used this storm exposure data to classify counties as exposed or unexposed
to four different hazards for each tropical cyclone and then explored patterns
over years with available data. Across the four storm hazards considered, there
was wide variation in the average number of county exposures per year
(Table~\ref{tab:exposuresummaries}). For tropical cyclone tornadoes,
there were on average about 50 county exposures per year within our study.
County exposures were more frequent for tropical cyclone wind exposures ($>$140
/ yr on average), even more frequent for tropical cyclone flood exposures
($>$200 / yr on average), and most frequent for tropical cyclone rain exposure
($>$400 / yr on average). For every hazard except tornadoes, we identified at
least one tropical cyclone that exposed over~250 counties
(Table~\ref{tab:exposuresummaries}).  However, the largest-extent tropical
cyclone varied across hazards: Frances in~2004 exposed the most counties based
on rain, Ike in~2008 based on wind, and Ivan in 2004 based on flooding and
tornadoes (Table~\ref{tab:exposuresummaries}).

When we calculated and mapped the average number of exposures per decade in
each county for single-hazard exposures (Figure~\ref{fig:averageexposure}),
strong geographical patterns were clear. Peak sustained wind exposure had a
strong coastal pattern, with almost all exposures in counties within about~200
\si{\kilo\metre} (124~mi) of the coastline. While tropical cyclone rain
exposure was also more frequent in coastal areas compared to inland areas,
there were also inland rain exposures in counties that were never classified as
exposed based on wind. Flood-based exposures were frequently in the
Mid-Atlantic region, with a pattern that skewed north compared with other
exposures. Rain and, to some extent, flood exposures were characterized by a
pattern defined by the Appalachian Mountains, with fewer exposures west of the
mountain range than to the east. Almost all tornado-based exposures were in
coastal states, with many in Florida, and almost none north of Maryland.
Patterns were similar when analysis was restricted to years with exposure data
for all four hazards available (1996\,--\,2011; Figure~S3).

\subsection*{Agreement across exposure metrics}

Finally, we measured the agreement between exposure classifications for each pair
of hazards, as well as with a proxy measurement based on
distance from the storm's track. We found that these exposure classifications
typically did not agree strongly between pairs of metrics---the set of counties
identified as ``exposed'' based on one metric often overlapped little with the set
identified as ``exposed'' by another metric, as storms frequently brought
different hazards to different locations.

Figure~\ref{fig:ivanexposure} shows as an example Hurricane Ivan in 2004.
For the distance-based metric, the counties assessed as exposed follow the
tropical cyclone's track. For the wind-based metric,
only counties near the tropical cyclone's first landfall were assessed as
exposed. For rain- and flood-based metrics, however, exposure extended to the
left of the track, including counties as far north as New York and Connecticut,
while for the tornado metric, exposed counties tended to be to the right of the
track and included several counties in central North Carolina, South Carolina,
and Georgia that were not identified as exposed to Ivan based on any other
metric. Figure~S4 provides similar maps for four other example tropical
cyclones (selected because they exposed many \ac{US}  counties based on at
least one metric).

We drew similar conclusions when we investigated all~46 tropical cyclones
between~1996 and~2011 (when data for all five metrics were
available) for which~100 or more counties were exposed based on at least one
metric (Figure~\ref{fig:jaccard}). In this figure, each row provides results
for one tropical cyclone, and each box in that row shows the Jaccard
coefficient for a pair of metrics.  For almost all tropical cyclones, Jaccard
coefficients were~$\le0.6$ between the distance-based metric and any of the
four hazard-based metrics (i.e., out of the counties assessed as exposed by at
least one of the two metrics in the pair,~60\si{\percent} or fewer were
assessed the same under both metrics). For many tropical cyclones, Jaccard
coefficients were~$\le0.5$, indicating at best moderate agreement.  Among the
four hazard-based metrics, the tornado-based metric had universally poor
agreement with other metrics in county-level classification across the tropical
cyclones considered.  Agreement between other pairs of metrics were better, but
still typically at best moderate (Jaccard coefficients usually $\le0.5$).

There were, however, exceptions\,---\,tropical cyclones in which exposure
assessment agreed well across several of the metrics considered.  For Floyd
in~1999 (Figure~S4) and Irene in~2011, for example, county-level classification
agreed moderately to well for all pairs of exposure metrics except those
involving the tornado-based metric (Figure~\ref{fig:jaccard}). For another set
of tropical cyclones (e.g., Ernesto in~2006, and Bertha in~1996, Isabel
in~2003), there was moderate to good agreement for all pairwise combinations of
distance, rain, and wind, but poor agreement for other combinations of metrics,
while for another set of storms (e.g., Matthew in~2004 and Katrina in~2005),
there was moderate to good agreement between distance and rain.