### 2020-02-06 tables/stormseverity.tex
Update based on latest hurricaneexposuredata version

% latex table generated in R 3.6.2 by xtable 1.8-4 package
% Thu Feb  6 11:53:51 2020
\begin{table}[ht]
\centering
\caption{Summary statistics for the number of county tropical cyclone exposures under each metric.} 
\label{tab:exposuresummaries}
\begin{tabular}{p{4.5cm}p{4.5cm}p{4.5cm}}
  \toprule
Metric (years available) & Mean (interquartile range) of county exposures per year & Tropical cyclone with most counties exposures (\# exposed counties) \\ 
  \midrule
Distance (1988--2015) & 369 (72, 466) & Beryl, 1994 (344) \\ 
  Rain (1988--2015) & 298 (48, 441) & Frances, 2004 (464) \\ 
  Wind (1988--2011) & 151 (55, 246) & Ike, 2008 (248) \\ 
  Flood (1996--2015) & 189 (67, 228) & Ivan, 2004 (317) \\ 
  Tornado (1996--2015) & 48 (15, 43) & Ivan, 2004 (91) \\ 
   \bottomrule
\end{tabular}
\end{table}