### 2020-02-19 tables/stormseverity.tex
Update year for storm severity table

% latex table generated in R 3.6.2 by xtable 1.8-4 package
% Wed Feb 19 10:41:15 2020
\begin{table}[ht]
\centering
\caption{Summary statistics for the number of county tropical cyclone exposures under each metric.} 
\label{tab:exposuresummaries}
\begin{tabular}{p{4.5cm}p{4.5cm}p{4.5cm}}
  \toprule
Metric (years available) & Mean (interquartile range) of county exposures per year & Tropical cyclone with most counties exposures (\# exposed counties) \\ 
  \midrule
Distance (1988--2018) & 399 (78, 503) & Beryl, 1994 (344) \\ 
  Rain (1988--2018) & 298 (48, 441) & Frances, 2004 (464) \\ 
  Wind (1988--2011) & 162 (55, 258) & Michael, 2018 (260) \\ 
  Flood (1996--2018) & 197 (76, 241) & Ivan, 2004 (317) \\ 
  Tornado (1988--2018) & 38 (8, 42) & Ivan, 2004 (91) \\ 
   \bottomrule
\end{tabular}
\end{table}