### 2020-02-19 writing/ehp_cover.tex

\title{Assessing United States county-level exposure for research on tropical cyclones and human health}

\author[1,*]{G.~Brooke Anderson} 
\author[1,2]{Joshua Ferreri}
\author[3]{Mohammad Al-Hamdan} 
\author[3]{William Crosson} 
\author[4]{Andrea Schumacher} 
\author[5]{Seth Guikema} 
\author[6]{Steven Quiring} 
\author[7]{Dirk Eddelbuettel} 
\author[1,8]{Meilin Yan} 
\author[9]{Roger D.~Peng}

\affil[1]{Department of Environmental \& Radiological Health Sciences, Colorado 
  State University, Fort Collins, CO, 80523} 
\affil[2]{University of Colorado Anschutz Medical Campus, School of Medicine, Aurora, CO, 80045} 
\affil[3]{Universities Space Research Association, NASA Marshall Space Flight 
  Center, Huntsville, AL, 35805}
\affil[4]{Cooperative Institute for Research in the Atmosphere, Colorado State
  University, Fort Collins, CO, 80523} 
\affil[5]{Department of Industrial and Operations Engineering, University of 
  Michigan, Ann Arbor, MI, 48109}
\affil[6]{Department of Geography, Ohio State University, Columbus, OH, 43210}
\affil[7]{Department of Statistics, University of
  Illinois at Urbana-Champaign, Champaign, IL, 61820} 
\affil[8]{Beijing Innovation Center for Engineering Science and Advanced Technology 
  and State Key Joint Laboratory of Environmental Simulation and Pollution Control, 
  Peking University, Beijing, China, 100871}
\affil[9]{Department of Biostatistics, Johns Hopkins Bloomberg School of Public 
  Health, Baltimore, MD, 21205}
\affil[*]{Corresponding author: G.~Brooke Anderson, brooke.anderson@colostate.edu,
  1681 Campus Delivery, Fort Collins, CO, 80523}

\date{}