### 2020-09-02 ehp_revision/tables/wind_misclass.tex

% latex table generated in R 4.0.2 by xtable 1.8-4 package
% Tue Sep  1 17:11:57 2020
\begin{table}[ht]
\centering
\caption{Caption. Limited to storms with at least 200 counties assessed as exposed based on at least one exposure metric considered in this study. Numbers are out of 2,396 counties in the study area (states in the eastern half of the US). Exposure assessment is based on the thresholds given in Table 1 of the main text. The Jaccard index shown in Figure [x] of the main text is based on numbers in the second through fourth columns (the value in the second column divide by the sum of numbers in the second through fourth columns). Storms are ordered based on the number of counties exposed to at least one of these two exposure metrics.} 
\begin{tabular}{cp{3cm}p{3cm}p{3cm}p{3cm}}
  \toprule
Storm & Classified as exposed based on both distance metric and wind metric & Classified as exposed based on distance metric but unexposed by wind metric & Classified as exposed based on wind metric but unexposed by distance metric & Classified as unexposed based on both distance metric and wind metric \\ 
  \midrule
Frances-2004 &  44 & 277 &  13 & 2062 \\ 
  Cindy-2005 &  17 & 304 &   4 & 2071 \\ 
  Dennis-2005 &  18 & 274 &  26 & 2078 \\ 
  Ike-2008 & 166 &  67 &  82 & 2081 \\ 
  Ivan-2004 &  29 & 255 &  27 & 2085 \\ 
  Jeanne-2004 &  41 & 256 &   7 & 2092 \\ 
  Allison-2001 &  21 & 266 &   0 & 2109 \\ 
  Isidore-2002 &  29 & 239 &   7 & 2121 \\ 
  Katrina-2005 &  66 & 175 &  32 & 2123 \\ 
  Gordon-2000 &  11 & 244 &   6 & 2135 \\ 
  Fay-2008 &  54 & 198 &   7 & 2137 \\ 
  Gustav-2008 &  36 & 198 &  23 & 2139 \\ 
  Bertha-1996 & 179 &   4 &  62 & 2151 \\ 
  Danny-1997 &  45 & 184 &  12 & 2155 \\ 
  Arlene-2005 &   4 & 232 &   0 & 2160 \\ 
  Bill-2003 &  19 & 211 &   0 & 2166 \\ 
  Dennis-1999 &  17 & 183 &  13 & 2183 \\ 
  Hanna-2008 & 135 &  57 &  12 & 2192 \\ 
  Isabel-2003 & 105 &  42 &  56 & 2193 \\ 
  Ernesto-2006 &  93 &  99 &   4 & 2200 \\ 
  Helene-2000 &  18 & 174 &   0 & 2204 \\ 
  Rita-2005 &  26 & 139 &  20 & 2211 \\ 
  Fran-1996 &  46 & 105 &  30 & 2215 \\ 
  Earl-1998 & 133 &  36 &  11 & 2216 \\ 
  Floyd-1999 & 124 &  15 &  40 & 2217 \\ 
  Irene-2011 & 114 &   6 &  46 & 2230 \\ 
  Lee-2011 &  34 & 110 &   0 & 2252 \\ 
  Josephine-1996 &  62 &  61 &   3 & 2270 \\ 
  Hermine-2010 &  22 & 101 &   1 & 2272 \\ 
  Bonnie-2004 &   0 & 109 &   0 & 2287 \\ 
  Frances-1998 &  15 &  47 &   0 & 2334 \\ 
   \bottomrule
\end{tabular}
\end{table}