### 2020-09-18 ehp_revision/tables/rain_misclass.tex

% latex table generated in R 4.0.2 by xtable 1.8-4 package
% Fri Sep 18 10:32:05 2020
\begin{table}[ht]
\centering
\caption{Agreement between rain-based exposure assessment and a distance-based
         proxy of storm exposure for tropical cyclones with at least 200 counties assessed 
         as exposed based on at least one exposure metric considered in this study. Numbers 
         are out of 2,396 counties in the study area (states in the eastern half of the US; 
         Figure 1 of the main text). Exposure assessment is based on the thresholds given in 
         Table 1 of the main text. The Jaccard index shown in Figure 7 of the main text is 
         calculated as the value in the second column divided by the sum of numbers in the 
         second through fourth columns.  Storms are ordered based on the number of counties 
         assessed as exposed to at least one of these two exposure metrics.).} 
\label{tab:misclassrain}
\begin{tabular}{cP{3cm}P{3cm}P{3cm}P{3cm}}
  \toprule
Storm & Exposed for both distance metric and rain metric & Exposed for distance metric but unexposed for rain metric & Exposed for rain metric but unexposed for distance metric & Unexposed for both distance metric and rain metric \\ 
  \midrule
Frances (2004) & 207 & 114 & 257 & 1,818 \\ 
  Ivan (2004) & 124 & 160 & 288 & 1,824 \\ 
  Isidore (2002) & 134 & 134 & 186 & 1,942 \\ 
  Ike (2008) & 113 & 120 & 213 & 1,950 \\ 
  Dennis (2005) & 48 & 244 & 124 & 1,980 \\ 
  Fay (2008) & 72 & 180 & 163 & 1,981 \\ 
  Jeanne (2004) & 144 & 153 & 113 & 1,986 \\ 
  Lee (2011) & 106 & 38 & 230 & 2,022 \\ 
  Gustav (2008) & 126 & 108 & 137 & 2,025 \\ 
  Allison (2001) & 106 & 181 & 67 & 2,042 \\ 
  Bill (2003) & 110 & 120 & 117 & 2,049 \\ 
  Cindy (2005) & 78 & 243 & 11 & 2,064 \\ 
  Floyd (1999) & 129 & 10 & 170 & 2,087 \\ 
  Katrina (2005) & 195 & 46 & 58 & 2,097 \\ 
  Danny (1997) & 82 & 147 & 54 & 2,113 \\ 
  Rita (2005) & 64 & 101 & 103 & 2,128 \\ 
  Gordon (2000) & 5 & 250 & 7 & 2,134 \\ 
  Arlene (2005) & 39 & 197 & 23 & 2,137 \\ 
  Dennis (1999) & 90 & 110 & 56 & 2,140 \\ 
  Ernesto (2006) & 124 & 68 & 63 & 2,141 \\ 
  Irene (2011) & 119 & 1 & 106 & 2,170 \\ 
  Hanna (2008) & 100 & 92 & 33 & 2,171 \\ 
  Hermine (2010) & 44 & 79 & 102 & 2,171 \\ 
  Fran (1996) & 89 & 62 & 73 & 2,172 \\ 
  Earl (1998) & 72 & 97 & 38 & 2,189 \\ 
  Bertha (1996) & 103 & 80 & 20 & 2,193 \\ 
  Helene (2000) & 28 & 164 & 11 & 2,193 \\ 
  Frances (1998) & 27 & 35 & 139 & 2,195 \\ 
  Josephine (1996) & 94 & 29 & 77 & 2,196 \\ 
  Bonnie (2004) & 56 & 53 & 56 & 2,231 \\ 
  Isabel (2003) & 88 & 59 & 17 & 2,232 \\ 
   \bottomrule
\end{tabular}
\end{table}